\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=2.5cm}

\title{Empirical Evidence of Theta-Field Predictivity within the Unified Biquaternion Theory}
\author{Ing.~David~Jaroš}
\date{\today}

\begin{document}
\maketitle

\section*{Abstract}
Recent experiments with Jacobi theta-based feature representations of market data
show a consistent but weak predictive signal when regressions are performed
on future price deltas.  
This paper interprets the results in the framework of the \emph{Unified Biquaternion Theory} (UBT)
and the \emph{Complex Consciousness Theory} (CCT), linking empirical findings
with the theoretical model of complex time and distributed consciousness.
We further outline implications for algorithmic trading systems and propose
directions for future development.

\section{Empirical Findings}
Across multiple causally clean experiments (ridge regression on theta bases),
correlations between predicted and realized deltas remain in the range
$R \approx 0.1$--$0.2$, with hit rates between 50--56\%.
The signal is statistically significant yet weak, indicating
that the theta decomposition captures latent periodic or toroidal
structures of market dynamics.

\section{Interpretation within UBT}
In UBT, every physical or informational system is a projection
of a higher-dimensional biquaternionic field:
\[
\Theta(q,\tau), \qquad \tau = t + i\psi,
\]
where the scalar part of $\tau$ corresponds to classical time
and the imaginary vector part $\psi$ encodes the phase or conscious direction.

The theta functions $\theta_k(\tau)$ represent toroidal harmonics
of this field.  
A market is thus modeled as a superposition of many such $\psi$-states,
each corresponding to an individual agent’s or process’s conscious trajectory.
The observable price series is an average over these projections.

\section{Why the Predictive Signal Is Weak}
The weak but persistent correlation arises because:
\begin{itemize}
  \item Individual theta components carry strong local coherence,
        but they are \emph{not orthogonal} and interfere destructively
        when aggregated across many conscious trajectories.
  \item The market average thus contains only the residual coherent part
        of the global phase---a weak, chaotic attractor.
  \item Ridge or Kalman filters act as projectors that isolate this coherent subspace.
\end{itemize}
In the language of CCT, the market behaves as a partially conscious system:
many $\psi$-vectors exist, but only their small synchronized fraction leaves
an imprint in physical time.

\section{Implications for Trading Agents}
For algorithmic applications, the key insight is that
theta-based representations contain real predictive information
only after an appropriate projection is learned.
Thus, a trading bot should:
\begin{enumerate}
  \item maintain an adaptive projection operator (e.g., Ridge or Kalman filter),
  \item restrict fitting to causal past data (walk-forward mode),
  \item use confidence weighting (e.g., $z$-score thresholding)
        to trade only when the projected coherence is strong.
\end{enumerate}
This corresponds, in UBT terms, to tuning the bot’s internal
\emph{conscious phase alignment} with the market’s collective $\psi$-field.

\section{Future Development}
\subsection*{Orthonormal Theta Basis}
To eliminate destructive interference, construct an orthonormal set
of theta components via a Gram--Schmidt process
over $\{\theta_0,\ldots,\theta_4\}$ evaluated along the real axis.
Analytic orthogonalization is preferable, but numerical vectors may suffice.

\subsection*{Biquaternionic Phase Expansion}
Extend the phase $\psi$ from a scalar to a full biquaternion vector:
\[
\tau = t + \mathbf{i}\psi_1 + \mathbf{j}\psi_2 + \mathbf{k}\psi_3,
\]
allowing multiple imaginary axes.
This increases the dimensionality of the latent space and may capture
multi-agent coherence patterns.

\subsection*{Advanced Filters}
Compare Ridge with alternative projectors:
Lasso, ElasticNet, and state-space Kalman filters for time-varying coefficients.
Ensure causal structure and validate using shuffle and out-of-sample tests.

\subsection*{Energy and Coherence Metrics}
Define an energy functional
\[
E_\Theta = \|\Theta\|^2_{\text{coh}} =
\sum_k w_k \langle \theta_k, \theta_k \rangle_{\text{coh}},
\]
to quantify the degree of market coherence.  
Higher $E_\Theta$ implies stronger alignment of conscious phases
and potentially higher predictive power.

\section{Conclusion}
The experiments confirm that Jacobi theta components contain
a measurable yet modest predictive edge, consistent with UBT’s
interpretation of markets as projections of a complex-time toroidal field.
Future work will focus on orthogonalization, dynamic filters, and
biquaternionic phase modeling to enhance coherence extraction
and enable a fully adaptive theta-based trading agent.

\end{document}

